---
title: "DOX_prot_Fig2_Network"
author: "Omar Johnson"
date: "2024-12-08"
output: html_document
---

# Load Libraries 
```{r Libraries, include=FALSE}
library(EDASeq)
library(RUVSeq)
library(RColorBrewer)
library(edgeR)
library(limma)
library(Biobase)
library(SummarizedExperiment)
library(tidyverse) 
library(ggfortify)
library(cluster)
library(edgeR)
library(limma)
library(BiocParallel)
library(qvalue)
library(pheatmap)
library(clusterProfiler)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(RColorBrewer)
library(DOSE)
library(UpSetR)
library(biomaRt)
library(ggridges)
library(reshape2)
library(BioNERO)
library(WGCNA)
library(impute)
library(dynamicTreeCut)
library(igraph)
library(ggraph)
```


# Read in Data 
```{r Read in data, include=FALSE}

# 1. RUVg Corrected data across all 10 samples that has been log2 transformed. It has not yet been quantile normalized. 
 RUVg_Log2_quantnormalized_all10samples <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Data_Frames/DIA_proteins/RUVIII_Imputed/RUVIII_10samples_log2_notquantilenormalized.csv", header = TRUE, row.names = 1)
RUVg_Log2_quantnormalized_all10samples <- RUVg_Log2_quantnormalized_all10samples^2
 

# 2. Toptable correspondding to the Diff.Abundance test from #1 
toptable_summary <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Data_Frames/DIA_proteins/RUVIII_Imputed/Toptable_summary_RUVIII.csv", header = TRUE, row.names = 1)


# 3. Meta data for all 10 samples in our study. 
Meta <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Data_Frames/DIA_proteins/RUVIII_Imputed/Meta.csv", header = TRUE, row.names = 1)



Toptable_Modules <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/Supplement/SUPP_Table_2.csv", header = TRUE)
Toptable_Modules %>% dim()

New_RNA_PRO_DF_2 <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Data_Frames/DIA_proteins/RUVIII_Imputed/New_RNA_PRO_DF.csv", header = TRUE, row.names = 1)
New_RNA_PRO_DF_2 %>% dim()

New_RNA_PRO_DF_3 <- merge(Toptable_Modules, New_RNA_PRO_DF_2, by.x = "Proteins", by.y = "uniprotswissprot")
New_RNA_PRO_DF_3$padjBHrev <- p.adjust(New_RNA_PRO_DF_3$P.Value.x, "BH")
New_RNA_PRO_DF_3



New_RNA_PRO_DF_3[New_RNA_PRO_DF_3$padjBHrev < 0.05, ]$hgnc_symbol
New_RNA_PRO_DF_3[New_RNA_PRO_DF_3$padjBHrev < 0.05, ] %>% dim()

hubs <- read.csv( file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/Supplement/hubs.csv", header = TRUE)

perform_module_comparisons_mutexc_2 <- function(df, module_col, value_col) {
  # Ensure the necessary columns exist
  if (!(module_col %in% names(df) && value_col %in% names(df))) {
    stop("Specified columns do not exist in the dataframe.")
  }

  # Get a list of all unique modules
  modules <- unique(df[[module_col]])

  # Initialize an empty list to store combined data frames
  combined_df_list <- list()

  # Initialize an empty dataframe to store results
  results <- data.frame(Module1 = character(),
                        Module2 = character(),
                        WilcoxPValue = numeric(),
                        stringsAsFactors = FALSE)

  # Loop through each module
  for (module in modules) {
    # Data for the current module
    current_data <- df %>% filter(!!sym(module_col) == module) %>% 
                    mutate(Group = as.character(module))

    # Data for all other modules
    other_data <- df %>% filter(!!sym(module_col) != module) %>% 
                    mutate(Group = paste("Not", module, sep=""))

    # Combine current module data with other module data
    combined_data <- rbind(current_data, other_data)

    # Add the combined data to the list
    combined_df_list[[module]] <- combined_data

    # Perform the Wilcoxon test
    test_result <- wilcox.test(current_data[[value_col]], other_data[[value_col]])

    # Add the results to the dataframe
    results <- rbind(results, data.frame(Module1 = module,
                                         Module2 = "Others",
                                         WilcoxPValue = test_result$p.value))
  }

  return(list("results" = results, "combined_data" = combined_df_list))
}

PPI_filepath <- "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/PPI_Functionally_annotated/Physical_PPI_no_weight.csv"


Toptable_summary_withadjP <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Data_Frames/Important_DF_Final/Toptable_summary_withadjP.csv", header =TRUE )

Toptable_summary_withadjP$padjBHrev <- p.adjust(Toptable_summary_withadjP$P.Value, "fdr")
Toptable_summary_withadjP$Is_DA_at_adj.p <- ifelse(Toptable_summary_withadjP$padjBHrev < 0.05, 1,0)
 Toptable_summary_withadjP[Toptable_summary_withadjP$Protein == "P78347", ]
# write.csv(x = Toptable_summary_withadjP,  file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Data_Frames/Important_DF_Final/Toptable_summary_iScience.csv")
 
 
 
 
 perform_fisher_test_FP <- function(vec1, vec2, vec1_name, vec2_name, plot = FALSE) {
  # Create labeled factors for vec1 and vec2
  vec1_label <- factor(vec1, labels = c(paste0("Not", vec1_name), paste0("Is", vec1_name)))
  vec2_label <- factor(vec2, labels = c(paste0("Not", vec2_name), paste0("Is", vec2_name)))

  # Create contingency table with labeled factors
  table <- table(vec1_label, vec2_label)

  # Perform Fisher's exact test
  test_result <- fisher.test(table)
  p_value <- test_result$p.value
OR <- test_result$estimate
CI <- test_result$conf.int

  # Prepare result
  result <- list(
    ContingencyTable = table,
    PValue = p_value, 
    Odds_ratio = test_result$estimate,
    Confidence_Interval = test_result$conf.int
  )

  # Generate plot 
  if (plot) {
    # Convert table to data frame for ggplot
    table_df <- as.data.frame(as.table(table))
    colnames(table_df) <- c("vec1_label", "vec2_label", "Freq")

    # Calculate totals for each vec1_label
    totals <- aggregate(Freq ~ vec1_label, data = table_df, sum)

    # Merge totals with table_df and calculate percentages
    table_df <- merge(table_df, totals, by = "vec1_label", all.x = TRUE)
    table_df$Percentage <- with(table_df, Freq.x / Freq.y * 100)
    table_df$Group <- table_df$vec2_label

    # Stacked bar chart
    p <- ggplot(table_df, aes(x = vec1_label, y = Percentage, fill = Group)) +
      geom_bar(stat = "identity", position = "stack") +  # Adjust position to "stack"
      facet_wrap(~ vec1_label) +
      theme_minimal() +
      labs(x = vec1_name, y = "Percentage", fill = vec2_name, title = paste("")) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

    result$Plot <- p
  }

  return(result)
 }
 
 
 
 
 
 
 TS_DF <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/Tissue_spec/Tissue_spec.csv", header = TRUE)
TS_DF


pLI_DF <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/pLI/Genes_pLI_Scores.csv", header = TRUE)







GO_enrichment_analysis_2 <- function(data_frame, 
                                   module_col, 
                                   gene_col, 
                                   ontology_categories, 
                                   modules_to_analyze) {
  
  # Initialize an empty list to store enrichment results as tibbles
  enrichment_results <- list()
  
  # Loop over each module and ontology category
  for (module in modules_to_analyze) {
    
    # Get unique genes in the specified module
    module_genes <- unique(data_frame[data_frame[[module_col]] == module, gene_col])
    
    # Define the unique background genes from the entire network
    background_genes <- unique(data_frame[[gene_col]])
    
    for (ont in ontology_categories) {
      # Perform GO Enrichment
      go_enrichment <- enrichGO(gene = module_genes,
                                OrgDb = org.Hs.eg.db,
                                keyType = "SYMBOL",
                                universe = background_genes,
                                ont = ont,
                                pvalueCutoff = 1,
                                qvalueCutoff = 1)
      
      # Convert the enrichment result to a tibble and add metadata columns
      go_enrichment_tibble <- as_tibble(go_enrichment) %>%
        mutate(Module = module, Ontology = ont)
      
      # Store the result in the list
      enrichment_results[[paste(module, ont, sep = "_")]] <- go_enrichment_tibble
      
      # Plotting the results
      p <- barplot(go_enrichment, showCategory = 10) +
        ggtitle(paste("Module:", module, "- Ontology:", ont))
      
      print(p)
    }
  }
  
  # Combine all tibbles into one data frame for easy access
  combined_results <- bind_rows(enrichment_results)
  
  return(combined_results)
}

New_RNA_PRO_DF_4 <-New_RNA_PRO_DF_3
New_RNA_PRO_DF_4TS <- merge(x =New_RNA_PRO_DF_4 , y = TS_DF, by.x = "Proteins", by.y = "Uniprot")
```


# Generate and view initial objects
```{r Create summarized experiment object}
# Undo log tranformation
RUVg_Log2_quantnormalized_all10samples %>% head()
RUVg_Log2_quantnormalized_all10samples %>% dim()


# Create summarized experiment object 
WGCNA_DF <- SummarizedExperiment(assays = as.matrix(RUVg_Log2_quantnormalized_all10samples))
dim(RUVg_Log2_quantnormalized_all10samples) 
exp_filt <- filter_by_variance(WGCNA_DF, n = 4178)
exp_filt %>% assay() %>% head()


toptable_summary %>% head()
Meta
```


# S3-A&B-SFT power threshold  
```{r SFT power threshold  }
# Determine the soft power threshold to get scale free shape
sft <- SFT_fit(exp_filt, net_type = "signed", rsquared = 0.7,  cor_method = "pearson")
sft$plot

```


# Modify workflow to construct a simpler network  
```{r  Modify workflow to have a simpler network}

exp2gcn2_EDIT <- function (exp, net_type = "signed", module_merging_threshold = 0.8, 
    SFTpower = NULL, cor_method = "spearman", TOM_type = NULL, 
    return_cormat = TRUE, verbose = FALSE) 
{
    params <- list(net_type = net_type, module_merging_threshold = module_merging_threshold, 
        SFTpower = SFTpower, cor_method = cor_method)
    exp <-  BioNERO:::handleSE(exp)
    if (is.null(SFTpower)) {
        stop("Please, specify the SFT power.")
    }
    if (verbose) {
        message("Calculating adjacency matrix...")
    }
    cor_matrix <- NULL
    if (return_cormat) {
        cor_matrix <- BioNERO::exp2cor(exp, cor_method = cor_method)
        adj_matrix <- BioNERO::cor2adj(cor_matrix, beta = SFTpower, net_type = net_type)
    }
    else {
        adj_matrix <- BioNERO::cor2adj(exp2cor(exp, cor_method = cor_method), 
            beta = SFTpower, net_type = net_type)
    }
    if (verbose) {
        message("Calculating topological overlap matrix (TOM)...")
    }
   tomtype <- net_type
  TOM <- WGCNA::TOMsimilarity(adj_matrix, TOMType = tomtype)
    geneTree <- hclust(as.dist(1 - TOM), method = "average")
    geneTree$height <- round(geneTree$height, 7)
    if (verbose) {
        message("Detecting coexpression modules...")
    }
    original_mods <- dynamicTreeCut::cutreeDynamicTree(dendro = geneTree, maxTreeHeight = 3,
        minModuleSize = 40, deepSplit = FALSE)
    nmod <- length(unique(original_mods))
    palette <- rev(WGCNA::standardColors(nmod))
    original_colors <- WGCNA::labels2colors(original_mods, colorSeq = palette)
    if (verbose) {
        message("Calculating module eigengenes (MEs)...")
    }
    me_list <- WGCNA::moduleEigengenes(t(exp), colors = original_colors, 
        softPower = SFTpower)
    me <- me_list$eigengenes
    original_metree <- hclust(as.dist(1 - cor(me, method = "spearman")), 
        method = "average")
    if (verbose) {
        message("Merging similar modules...")
    }
    merged <- BioNERO:::merge_modules(exp, original_colors, me, palette, 
        dissimilarity = 1 - module_merging_threshold, cor_method = cor_method)
    new_colors <- merged$colors
    new_mes <- merged$newMEs
    genes_modules <- data.frame(Genes = rownames(exp), Modules = new_colors)
    if (verbose) {
        message("Calculating intramodular connectivity...")
    }
    kwithin <- WGCNA::intramodularConnectivity(adj_matrix, new_colors)
    result_list <- list(adjacency_matrix = adj_matrix, MEs = new_mes, 
        genes_and_modules = genes_modules, kIN = kwithin, correlation_matrix = cor_matrix, 
        params = params, dendro_plot_objects = list(tree = geneTree, 
            Unmerged = original_colors, Merged = new_colors))
    return(result_list)
}
```


# Fig S3 
```{r Generate network with max merge threshold }
net <- exp2gcn2_EDIT(
  exp_filt, net_type = "signed", module_merging_threshold = 0.9999, SFTpower = 20, 
  cor_method = "pearson")

# Eigengene networks
WGCNA::plotEigengeneNetworks(net$MEs, "", marDendro = c(3, 5, 2, 6), plotHeatmaps = FALSE)


WGCNA::plotEigengeneNetworks(net$MEs, "", marHeatmap = c(3, 4, 2, 2), plotDendrograms = FALSE)



# Genes per module 
plot_ngenes_per_module(net)
```


# Figure-2-A & Fig S3 Weighted protein co-expression network
```{r Generate network with desired merge threshold}

net <- exp2gcn2_EDIT(
  exp_filt, net_type = "signed", module_merging_threshold = 0.85, SFTpower = 20, 
  cor_method = "pearson")


net_unmerged <- exp2gcn2_EDIT(
  exp_filt, net_type = "signed", module_merging_threshold = 1, SFTpower = 20, 
  cor_method = "pearson")



# Eigengene networks
WGCNA::plotEigengeneNetworks(net$MEs, "", marDendro = c(3, 5, 2, 6), plotHeatmaps = FALSE, excludeGrey = TRUE)


WGCNA::plotEigengeneNetworks(net$MEs, "", marHeatmap = c(3, 4, 2, 2), plotDendrograms = FALSE)

# Genes per module 
plot_ngenes_per_module(net)
WGCNA::plotEigengeneNetworks(net_unmerged$MEs, "", marDendro = c(3, 5, 2, 6), plotHeatmaps = FALSE, excludeGrey = TRUE)

BioNERO::plot_eigengene_network(gcn = net)
BioNERO::plot_eigengene_network(gcn = net_unmerged)

```


# Fig-2-A-Define trait and get eigen proteins
```{r Define trat and get eigen proteins, fig.width = 8, fig.height = 3}

Meta

WGCNA_DF$trait <- c(1,1,1,1,1,0,0,0,0,0)
names(WGCNA_DF$trait) <- c("S1", "S3", "S5", "S7", "S9", "S2", "S4", "S6", "S8", "S10")


eigenmatrix <- net$MEs %>% as.matrix()
eigenmatrix


correlationResults <- cor(eigenmatrix, WGCNA_DF$trait, method= "pearson")

pvalues <- sapply(1:ncol(eigenmatrix), function(column) cor.test(eigenmatrix[,column], WGCNA_DF$trait)$p.value )

WGCNA_DF$trait <- as.matrix(WGCNA_DF$trait)

Modeigen <- as.matrix(net$MEs)


# Trait data frame
trait_data <- data.frame(
  ME = correlationResults %>% rownames(),
  Value = correlationResults[,1],
  PValue = pvalues
)



Drug_trait_results <- trait_data



WGCNA_DF$trait <- c(1,2,3,3,3,1,2,3,3,3)
names(WGCNA_DF$trait) <- c("S1", "S3", "S5", "S7", "S9", "S2", "S4", "S6", "S8", "S10")


eigenmatrix <- net$MEs %>% as.matrix()



correlationResults <- cor(eigenmatrix, WGCNA_DF$trait, method="pearson")

pvalues <- sapply(1:ncol(eigenmatrix), function(column) cor.test(eigenmatrix[,column], WGCNA_DF$trait)$p.value )

WGCNA_DF$trait <- as.matrix(WGCNA_DF$trait)

Modeigen <- as.matrix(net$MEs)



# Trait data frame
trait_data <- data.frame(
  ME = correlationResults %>% rownames(),
  Value = correlationResults[,1],
  PValue = pvalues
)
trait_data$neglogP <- -log(trait_data$PValue)


# Get modules that significantly associate with DOX treatment 
Signifigant_modules <-  trait_data[trait_data$PValue < 0.01, ]$ME



# Get your genes/proteins and modules DF 
Gene_mod_DF <- net$genes_and_modules 


Drug_trait_results_sub <- Drug_trait_results[,c(1,2,3)]
colnames(Drug_trait_results_sub) <- c("Module", "cor", "pval")
Drug_trait_results_sub$Trait <- "DOX"




Ind_trait_sub <- trait_data[,c(1,2,3)]
colnames(Ind_trait_sub) <- c("Module","cor", "pval")
Ind_trait_sub$Trait <- "IND."

Fulltrait_df <- rbind(Drug_trait_results_sub,Ind_trait_sub)



Fulltrait_df$annotation <- ifelse(Fulltrait_df$pval < 0.01, "*", "")
module_order <- c("MEbrown", "MEblack", "MEmagenta", "MElightgreen", "MEblue",      "MEgreen","MElightyellow","MEsalmon","MEyellow","MEcyan","MEdarkgreen", "MEmidnightblue", "MEgrey")  

# Factor the Module column in Fulltrait_df
Fulltrait_df$Module <- factor(Fulltrait_df$Module, levels = module_order)


ggplot(Fulltrait_df, aes(x = Module, y = Trait, fill = cor)) +
  geom_tile() +
  # First geom_text for annotations
  geom_text(aes(label = annotation), color = "black", size = 10) +
  # Second geom_text for correlation values
  geom_text(aes(label = sprintf("%.2f", cor)), color = "black", size = 4, vjust = .8) +
  scale_fill_gradient2(
    low = "blue", 
    high = "red", 
    mid = "white", 
    midpoint = 0, 
    name = "Correlation",
    limits = c(-1, 1)  # Set the limits for the color scale
  ) +
  labs(x = "Trait", y = "Module") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Using Figure-2-B - Pairwise differential abundance of proteins
```{r Volcano}

# Get your genes/proteins and modules DF 
Gene_mod_DF <- net$genes_and_modules 

# Module data gets merged with toptable
toptable_summary %>% head()

Toptable_Modules <- merge(toptable_summary, 
                          Gene_mod_DF, by.x = "Protein", by.y = "Genes")


New_RNA_PRO_DF_4$threshold_P_adj <- ifelse(New_RNA_PRO_DF_4$padjBHrev < 0.05, 1,0)
New_RNA_PRO_DF_4$threshold_P_adj <- New_RNA_PRO_DF_4$threshold_P_adj %>% as.character()
# Create the volcano plot
ggplot(New_RNA_PRO_DF_4)+
  geom_point(mapping = aes(x = Pro_LogFC, y = -log10(padjBHrev), color = threshold_P_adj))+
  xlab("log2FC")+
  ylab("-log10 adjusted p-value")+
  ylim(0, 3.5)+
  xlim(-3, 3)+
  theme(legend.position = "none", 
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25)))+
  theme_classic()

```



#Grey module enrichment
```{r }

New_RNA_PRO_DF_4 <-New_RNA_PRO_DF_3
New_RNA_PRO_DF_4$Is_grey <- ifelse(New_RNA_PRO_DF_4$Modules.x == "grey", 1, 0)




perform_fisher_test_FP(vec1 =New_RNA_PRO_DF_4$Is_grey , vec2 = New_RNA_PRO_DF_4$Is_pQTL, vec1_name = "", vec2_name = "")


Grey_pQTL <- perform_fisher_test_FP(vec1 =New_RNA_PRO_DF_4$Is_grey , vec2 = New_RNA_PRO_DF_4$Is_pQTL, vec1_name = "", vec2_name = "")
Grey_pQTL


Grey_Mut_Int. <- perform_fisher_test_FP(vec1 =New_RNA_PRO_DF_4$Is_grey , vec2 = New_RNA_PRO_DF_4$pLI_Mut.Intolerant, vec1_name = "", vec2_name = "")
Grey_Mut_Int.

Grey_Mut_Tol. <- perform_fisher_test_FP(vec1 =New_RNA_PRO_DF_4$Is_grey , vec2 = New_RNA_PRO_DF_4$pLI_Mut.Tolerant, vec1_name = "", vec2_name = "")
Grey_Mut_Tol.



FP_List <- list( Grey_pQTL, Grey_Mut_Int., Grey_Mut_Tol.)

FP_DF <- data.frame(
  Odds_ratio = numeric(length(FP_List)),
  Lower_CI = numeric(length(FP_List)),
  Upper_CI = numeric(length(FP_List)), 
  Pval = numeric(length(FP_List))
)

for (i in 1:length(FP_List)) {
  FP_DF$Odds_ratio[i] <- FP_List[[i]]$Odds_ratio
  FP_DF$Lower_CI[i] <- FP_List[[i]]$Confidence_Interval[1]
  FP_DF$Upper_CI[i] <- FP_List[[i]]$Confidence_Interval[2]
  FP_DF$Pval[i] <- FP_List[[i]]$PValue
}

# Add row names for the labels in the forest plot
FP_DF$Label <- c("Grey_pQTL", "Grey_Mut_Int.", 'Grey_Mut_Tol.')


FP_DF$Label <- factor(FP_DF$Label, levels = rev(c("Grey_pQTL", "Grey_Mut_Int.", 'Grey_Mut_Tol.')))

FP_DF
# Add a new column to indicate "cis" or "trans"
FP_DF$Type <- c("pQTL", "Mut.Int.", "Mut.Tol.")


# Plot with color based on "Type"
ggplot(FP_DF, aes(x = Label, y = Odds_ratio, ymin = Lower_CI, ymax = Upper_CI, color = Type)) +
  geom_pointrange() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  coord_flip() +
  labs(
    title = "Odds of mutation tolerance",
    x = "",
    y = "Odds Ratio (95% CI)"
  ) +
  theme_classic() 


```



# Fig S9-10 Using EnrichGO for all modules (See supplementary tables)
```{r }

# 
# 
# net_unmerged_DF_merged_mods <- New_RNA_PRO_DF_4 %>% unique()
# BP_DF <- GO_enrichment_analysis_2(data_frame = New_RNA_PRO_DF_4, 
#                        module_col ="Modules.x" , 
#                        gene_col = "hgnc_symbol", 
#                        ontology_categories = "BP", 
#                        modules_to_analyze = net_unmerged_DF_merged_mods)
# 
# 
# MF_DF <- GO_enrichment_analysis_2(data_frame = New_RNA_PRO_DF_4, 
#                        module_col ="Modules.x" , 
#                        gene_col = "hgnc_symbol", 
#                        ontology_categories = "MF", 
#                        modules_to_analyze = net_unmerged_DF_merged_mods)
# 
# 
# 
# CC_DF <- GO_enrichment_analysis_2(data_frame = New_RNA_PRO_DF_4, 
#                        module_col ="Modules.x" , 
#                        gene_col = "hgnc_symbol", 
#                        ontology_categories = "CC", 
#                        modules_to_analyze = net_unmerged_DF_merged_mods)



```


# Figure-2-D DOX-response effect size
```{r Ridge 1}

Toptable_Modules_DOXcorset <- Toptable_Modules[Toptable_Modules$Modules %in% c("green", "lightyellow" , "salmon", "darkgreen" ,"midnightblue"), ]



# Density 
ggplot(Toptable_Modules_DOXcorset, aes(x = logFC, fill = Modules)) +
  geom_density(alpha = 0.9) + 
  labs(title = "Distribution of logFC Values in DOX correlated modules",
       x = "logFC",
       y = "Density") +
  theme_minimal() +
  theme(legend.position = "right")+
  xlim(-2,2)+
  ylim(0, 4.5)+
  scale_fill_identity(guide = "legend")





Toptable_Modules_DOXcorset_SAYANversion <- Toptable_Modules_DOXcorset


Toptable_Modules_DOXcorset_SAYANversion$Modules[Toptable_Modules_DOXcorset_SAYANversion$Modules == "lightyellow"] <- "goldenrod" 


# Full graph 
module_order <- c("green","darkgreen","midnightblue", "salmon","goldenrod")

# Factor the Module column in Fulltrait_df
Toptable_Modules_DOXcorset_SAYANversion$Modules <- factor(Toptable_Modules_DOXcorset_SAYANversion$Modules, levels = rev(module_order))



# Density 
ggplot(Toptable_Modules_DOXcorset_SAYANversion, aes(x = logFC, fill = Modules)) +
  geom_density(alpha = 0.9) + 
  labs(title = "Distribution of logFC Values in DOX correlated modules",
       x = "logFC",
       y = "Density") +
  theme_classic() +
  theme(legend.position = "right")+
  xlim(-2,2)+
  ylim(0, 4)+
  scale_fill_identity(guide = "legend")


Toptable_Modules$Modules %>% unique()
Toptable_Modules_NonDOXcorset <- Toptable_Modules[Toptable_Modules$Modules %in% c("black", "cyan" ,"lightgreen", "magenta", "blue" ,"brown", "yellow"), ]


# Density
ggplot(Toptable_Modules_NonDOXcorset, aes(x = logFC, fill = Modules)) +
  geom_density(alpha = 0.9 ) + 
  labs(title = "Distribution of logFC Values in modules not correlaated to DOX",
       x = "logFC",
       y = "Density") +
  theme_minimal() +
  theme(legend.position = "right")+
  xlim(-2,2)+
  ylim(0, 4.5)+
  scale_fill_identity(guide = "legend")


```


```{r , include = FALSE}

Conserved_Mod_result_trait_data_DNAD_24_merged <- read.csv(file = "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/Tissue_spec/Conserved_Mod_result_trait_data_DNAD_24_merged.csv", header = TRUE, row.names = 1)

```

# TS DOX corr line 
```{r }

cor_test_result <- cor.test(Conserved_Mod_result_trait_data_DNAD_24_merged$conserved_proteins, Conserved_Mod_result_trait_data_DNAD_24_merged$cor)
cor_test_result

# Extract correlation coefficient and p-value
cor_coef <- cor_test_result$estimate
p_value <- cor_test_result$p.value

# Format the annotation text
annotation_text <- paste0("r = ", round(cor_coef, 2), 
                          "\nP-value = ", signif(p_value, 3))

# Plot with annotation
ggplot(Conserved_Mod_result_trait_data_DNAD_24_merged, aes(x = conserved_proteins, y = cor)) +
  geom_point(color = "blue", alpha = 0.5, size = 3) +  # Scatter plot points
  geom_smooth(method = "lm", color = "red", fill = "pink", se = TRUE) +  # Best fit line with confidence interval
  labs(
    title = "",
    x = "Tissue specificity for heart ventricle",
    y = "Correlation to DOX"
  ) +
  theme_classic() +
  annotate("text", label = annotation_text, color = "black", size = 1, hjust = 3)



ggplot(Conserved_Mod_result_trait_data_DNAD_24_merged, aes(x = conserved_proteins, y = cor, color = Modules.x)) +
  geom_point(alpha = 1, size = 3) +  # Scatter plot points with colors from Modules.x
  geom_smooth(method = "lm", color = "red", fill = "pink", se = TRUE) +  # Best fit line with confidence interval
  labs(
    title = "",
    x = "Tissue specificity for heart ventricle",
    y = "Correlation to DOX"
  ) +
  theme_classic() +
  annotate("text", x = max(Conserved_Mod_result_trait_data_DNAD_24_merged$percent_conserved) * 0.8, 
           y = max(abs(Conserved_Mod_result_trait_data_DNAD_24_merged$cor)) * 0.9,
           label = annotation_text, color = "black", size = 3, hjust = 0, vjust = -3) 


```

```{r, include=FALSE}

pLI_DF_File_path <- "/Users/omarjohnson/Documents/Projects/Dox_Proteomics/Data/Proteomics/Data_sets/pLI/Genes_pLI_Scores.csv"

library(vioplot)
```

# S14 pLI and individual corr. 
```{r, fig.width=10, fig.height=8 }

pLI_Data <- read.csv(file = pLI_DF_File_path, header = TRUE)

pLI_Data_sub <- merge(pLI_Data, New_RNA_PRO_DF_3, by.x = "gene", by.y = "hgnc_symbol")

pLI_Data_sub2 <- pLI_Data_sub


pLI_Data_sub2$Modules.y %>% unique()
DOXcorr_pLI <- pLI_Data_sub2[pLI_Data_sub2$Is_DOXcorrelated == 1, ]$pLI

Not_DOXcorr_Ind_pLI <- pLI_Data_sub2[(pLI_Data_sub2$Is_DOXcorrelated == 0) & (pLI_Data_sub2$Modules.x %in% c("darkred", "magenta", "royalblue")), ]$pLI

Not_DOXcorr_Not_Ind_pLI <- pLI_Data_sub2[(pLI_Data_sub2$Is_DOXcorrelated == 0) & (!pLI_Data_sub2$Modules.x %in% c("darkred", "magenta", "royalblue")), ]$pLI

DOXcorr_pLI %>% median(na.rm = TRUE)
Not_DOXcorr_Ind_pLI %>% median(na.rm = TRUE)
Not_DOXcorr_Not_Ind_pLI  %>% median(na.rm = TRUE)


# Create the boxplot
boxplot(DOXcorr_pLI, Not_DOXcorr_Not_Ind_pLI, Not_DOXcorr_Ind_pLI, names = c("+ DOXcorr - Ind corr.", "- DOXcorr. -Ind corr.", "- DOXcorr. + Ind corr"),
        
        main = "pLI across module types",
        xlab = "", ylab = "pLI")


vioplot(DOXcorr_pLI, Not_DOXcorr_Not_Ind_pLI, Not_DOXcorr_Ind_pLI, names = c("+ DOXcorr - Ind corr.", "- DOXcorr. -Ind corr.", "- DOXcorr. + Ind corr"),
        
        main = "pLI across module types",
        xlab = "", ylab = "pLI")

wilcox.test(DOXcorr_pLI, Not_DOXcorr_Ind_pLI)
wilcox.test(Not_DOXcorr_Not_Ind_pLI, Not_DOXcorr_Ind_pLI)
wilcox.test(DOXcorr_pLI, Not_DOXcorr_Not_Ind_pLI)

```


# Figure-2-E Hub protein expresssion examples 
```{r Ridge 2}
boxplot(as.matrix(log2(RUVg_Log2_quantnormalized_all10samples))["Q96GQ7",] ~ Meta$Cond, data=Meta, main="DDX27",
        ylab="Log2 Abundance", xlab="", col = c("green","green"), ylim = c(8,9))


boxplot(as.matrix(log2(RUVg_Log2_quantnormalized_all10samples))["Q9BRB3",] ~ Meta$Cond, data=Meta, main="PIGQ",
        ylab="Log2 Abundance", xlab="", col = c("yellow","yellow"),  ylim = c(8,9))

```


# Figure-2-F-Hub protein co-expression network 
```{r 2F, fig.width=20, fig.height=20}
#### Version1####
hubs_traitcorr <- hubs[hubs$Module %in% c( "darkgreen", "green" , "lightyellow",  "midnightblue", "salmon"), ]

hubs_traitcorr_toptable <-  merge(hubs_traitcorr, Toptable_Modules, by.x="Gene" , by.y= "Protein")


My_edge_list <- BioNERO::get_edge_list(net)

Toptable_Modules_2 <- Toptable_Modules
Toptable_Modules_2$adj.P.Val <- Toptable_Modules_2$P.Value %>% p.adjust("fdr")
Toptable_Modules_2[(Toptable_Modules_2$adj.P.Val < 0.05) & (Toptable_Modules_2$logFC > 0), ] %>% dim()

# Get a data frame of DOX corr. hubs connecting all to all 
My_edge_list_hubs <- My_edge_list %>%
  filter(Gene1 %in% hubs_traitcorr$Gene & Gene2 %in% hubs_traitcorr$Gene)
# looks like it is too big for plotting try something else 



# Create a helper dataframe with Gene and Module from hubs_traitcorr
helper_df <- hubs_traitcorr %>% dplyr::select(Gene, Module)

# Now, subset My_edge_list based on the conditions
My_edge_list_hubs_same_module <- My_edge_list_hubs %>%
  dplyr::filter(
    Gene1 %in% helper_df$Gene & 
    Gene2 %in% helper_df$Gene
  ) %>%
  dplyr::filter(
    helper_df$Module[match(Gene1, helper_df$Gene)] == 
    helper_df$Module[match(Gene2, helper_df$Gene)]
  )

My_edge_list_hubs_same_module %>% head()
My_edge_list_hubs_same_module_filt <- My_edge_list_hubs_same_module[My_edge_list_hubs_same_module$Weight > 0.9, ] 



#### Version 2 fully connected #### 
# Create an igraph graph object

graph_network <- graph_from_data_frame(d = My_edge_list_hubs[My_edge_list_hubs$Weight > 0.9, ], directed = FALSE)

# Prepare the node attribute data frame with gene and module information
node_attributes <- dplyr::select(hubs_traitcorr, Gene, Module) %>% unique()

# Add the module information to the graph
V(graph_network)$Module <- node_attributes$Module[match(names(V(graph_network)), node_attributes$Gene)]


ggraph(graph_network, layout = "stress") +
  geom_edge_link(aes(edge_color = Weight), show.legend = TRUE) +
  geom_node_point(aes(color = Module), stroke = 0.5) +  # Color nodes by module
  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
  scale_edge_color_continuous(low = "lightgrey", high = "black", limits = c(0.9, 1)) +  # Scale edge color by weight
  scale_color_manual(values = unique(hubs_traitcorr$Module)) +  # Set colors for each module
  theme_graph() +
  labs(color = "Module")

ggraph(graph_network, layout = "stress") +
  geom_edge_link(aes(edge_color = Weight, alpha = Weight), show.legend = TRUE) +  # Edge color and alpha by Weight
  geom_node_point(aes(color = Module), size = 2) +  # Color nodes by module, adjust node size
  geom_node_text(aes(label = name), repel = TRUE, size = 2) +  # Adjust text size
  scale_edge_color_continuous(low = "white", high = "black", limits = c(0.9, 1)) +  # Scale edge color by weight
  scale_color_manual(values = unique(hubs_traitcorr$Module)) +  # Set colors for each module
  theme_graph() +
  labs(color = "Module")



```

# Fig S6 Network hub percentages 
```{r Network hub percentages, fig.width=20, fig.height=20}
Hub_sum_table <- hubs %>% dplyr::group_by(Module) %>%
  dplyr::summarize(Number = n())

Hub_sum_table$prop <- (Hub_sum_table$Number)/sum(Hub_sum_table$Number)

Percent_Hub_Df <- data.frame(
  
  percent_hubs = c(1.24, 19.90, 3.23, 5.21, 0.99, 14.40, 10.90, 5.21, 1.49, 17.90, 13.90, 5.71), 
  
  Number = c(5,80,13,21,4,58,44,21,6,72,56,23),
  
  module_order = c("MEbrown", "MEdarkred", "MEmagenta", "MElightgreen", "MEblue",      "MEgreen","MElightyellow","MEsalmon","MEyellow","MEroyalblue","MEdarkgreen", "MEmidnightblue")
)

Percent_Hub_Df$prop.hub <- Percent_Hub_Df$percent_hubs/100
  Percent_Hub_Df$Number <- Percent_Hub_Df$Number %>% as.character()

# Factor the Module column in Fulltrait_df_2
Percent_Hub_Df$module_order <- factor(Percent_Hub_Df$module_order, levels = c("MEbrown", "MEdarkred", "MEmagenta", "MElightgreen", "MEblue",      "MEgreen","MElightyellow","MEsalmon","MEyellow","MEroyalblue","MEdarkgreen", "MEmidnightblue") )

Percent_Hub_Df$Number <- Percent_Hub_Df$Number %>% as.double()
Percent_Hub_Df$Module <- c("brown", "darkred", "magenta", "lightgreen", "blue",      "green","lightyellow","salmon","yellow","royalblue","darkgreen", "midnightblue")
  

Percent_Hub_Df$Module <- factor(x = Percent_Hub_Df$Module, levels = c("green","darkgreen","midnightblue","salmon", "lightyellow", "lightgreen", "blue","magenta", "darkred","brown", "yellow","royalblue", "grey"))
  
  
ggplot(Percent_Hub_Df, aes(x = Module, y = Number, fill = Module)) +
  geom_col()+
  scale_fill_identity()+
  theme_classic()
  
```

# Fig S6 Network protein connectivity comparison
```{r Network protein connectivity comparison}
DOXcorproteins <- Toptable_Modules[Toptable_Modules$Modules %in% c("green", "darkgreen", "midnightblue", "salmon", "lightyellow"), ]$Protein
KIN_DF <- net$kIN
KIN_DF$uniprot <- rownames(KIN_DF)
KIN_DF_TT <- merge( Toptable_Modules, KIN_DF, by.x = "Protein", by.y = "uniprot")


kIN_DF_Hubs <- KIN_DF_TT[KIN_DF_TT$Protein %in% hubs$Gene, ]
kIN_DF_Not_Hubs <- KIN_DF_TT[!KIN_DF_TT$Protein %in% hubs$Gene, ]
kIN_DF_Hubs_DOXcorr <- kIN_DF_Hubs[kIN_DF_Hubs$Protein %in% DOXcorproteins, ]
kIN_DF_Hubs_NOT_DOXcorr <- kIN_DF_Hubs[!kIN_DF_Hubs$Protein %in% DOXcorproteins, ]


wilcox.test(kIN_DF_Hubs$kWithin, kIN_DF_Not_Hubs$kWithin)
wilcox.test(kIN_DF_Hubs_DOXcorr$kWithin, kIN_DF_Hubs_NOT_DOXcorr$kWithin)
wilcox.test(kIN_DF_Hubs_NOT_DOXcorr$kWithin, kIN_DF_Not_Hubs$kWithin)


kIN_DF_Hubs$kWithin %>% length()
kIN_DF_Not_Hubs$kWithin %>% length()
kIN_DF_Hubs_DOXcorr$kWithin  %>% length()
kIN_DF_Hubs_NOT_DOXcorr$kWithin  %>% length()



# Do the boxplot 
# Create a data frame
pLI_Hubs_DF_Boxplot <- data.frame(
  values = c(kIN_DF_Hubs_DOXcorr$kWithin, kIN_DF_Hubs_NOT_DOXcorr$kWithin, kIN_DF_Not_Hubs$kWithin, kIN_DF_Hubs$kWithin ),
  group = factor(c(rep("DOX cor hub", length(kIN_DF_Hubs_DOXcorr$kWithin))
                   , rep("Non-DOX cor hub", length(kIN_DF_Hubs_NOT_DOXcorr$kWithin)), rep("Non-hub", length(kIN_DF_Not_Hubs$kWithin)), rep("All-hub", length(kIN_DF_Hubs$kWithin))))
)




# Create boxplot
ggplot(pLI_Hubs_DF_Boxplot, aes(x = group, y = values)) +
  geom_boxplot() +
  labs(x = "", y = "connectivity (kIN)") +
  theme_classic()+
  coord_cartesian(ylim = c(0,200))+
  theme(axis.text.x = element_text(face = "bold", color = "black",
size = 8, angle = 0, vjust = 0 ))
```

# Fig S15 Physical protein interactor co-expression
```{r }

PPI_Noweight<- read.csv(file = PPI_filepath, header = TRUE)
PPI_Confident <- PPI_Noweight[PPI_Noweight$stringdb..score >= 0.9, ]
PPI_Confident %>% dim()
Toptable_Modules_PPI <- Toptable_Modules




# Step 1: Join Toptable_Modules_PPI with PPI_Confident to find interactions for each protein
# We perform two joins to capture interactions where the protein is in either query_term1 or query_term2
interactions <- PPI_Confident %>%
  left_join(Toptable_Modules_PPI, by = c("query_term1" = "Protein")) %>%
  rename(Module1 = Modules) %>%
  left_join(Toptable_Modules_PPI, by = c("query_term2" = "Protein")) %>%
  rename(Module2 = Modules) %>%
  dplyr::select(query_term1, query_term2, Module1, Module2)

# Step 2: Calculate the proportion of interactors in the same module for each protein
module_overlap <- interactions %>%
  mutate(Same_Module = Module1 == Module2) %>%
  group_by(query_term1) %>%
  summarise(
    Total = n(),
    Same_Module_Count = sum(Same_Module, na.rm = TRUE),
    Prop_PPI_overlap = Same_Module_Count / Total
  ) %>%
  ungroup()

# Step 3: Merge the calculated proportions back into the Toptable_Modules_PPI
Toptable_Modules_PPI <- Toptable_Modules_PPI %>%
  left_join(module_overlap, by = c("Protein" = "query_term1"))

# Step 4: Ensure all proteins have a value, replacing NA with 0 where no interactors are found
Toptable_Modules_PPI$Prop_PPI_overlap <- ifelse(is.na(Toptable_Modules_PPI$Prop_PPI_overlap), 0, Toptable_Modules_PPI$Prop_PPI_overlap)

# View the results
print(Toptable_Modules_PPI)
Toptable_Modules_PPI$Prop_PPI_overlap %>% hist()



# For all module proteins 
ggplot(Toptable_Modules_PPI, aes(x = Prop_PPI_overlap, y = Modules, fill = Modules)) +
    geom_boxplot(alpha = 0.8) +
  geom_point(alpha = 0.3)+
  labs(title = "% module proteins with same mod PPI",
       x = "% module overlap",
       y = "") +
  theme_minimal()+
  scale_fill_identity() 




PPI_mod_res <- perform_module_comparisons_mutexc_2(df = Toptable_Modules_PPI , module_col = "Modules", value_col = "Prop_PPI_overlap")
PPI_mod_res$results[PPI_mod_res$results$WilcoxPValue < 0.05, ]


module_summary <- Toptable_Modules_PPI %>%
  group_by(Modules) %>%
  summarise(
    Mean_PPI_Overlap = mean(Prop_PPI_overlap, na.rm = TRUE),
    Median_PPI_Overlap = median(Prop_PPI_overlap, na.rm = TRUE)
  )
module_summary$Mean_PPI_Overlap %>% range()









# For hub proteins 
Toptable_Modules_PPI_hub <- Toptable_Modules_PPI[Toptable_Modules_PPI$Protein %in%hubs$Gene, ]

Toptable_Modules_PPI_NOThub <- Toptable_Modules_PPI[!Toptable_Modules_PPI$Protein %in%hubs$Gene, ]

wilcox.test(Toptable_Modules_PPI_hub$Prop_PPI_overlap, Toptable_Modules_PPI_NOThub$Prop_PPI_overlap)

pLI_Hubs_DF_Boxplot <- data.frame(
  values = c(Toptable_Modules_PPI_hub$Prop_PPI_overlap, Toptable_Modules_PPI_NOThub$Prop_PPI_overlap),
  group = factor(c(rep("Hubs", length(Toptable_Modules_PPI_hub$Prop_PPI_overlap)), rep("Not Hubs", length(Toptable_Modules_PPI_NOThub$Prop_PPI_overlap))))
)


# Create boxplot
ggplot(pLI_Hubs_DF_Boxplot, aes(x = group, y = values)) +
  geom_violin() +
  geom_point(alpha = 0.2)+
  labs(x = "", y = "% Physical PPI in saame module") +
  theme_minimal()

ggplot(pLI_Hubs_DF_Boxplot, aes(x = group, y = values)) +
  geom_boxplot() +
  geom_point(alpha = 0.2)+
  labs(x = "", y = "% Physical PPI in saame module") +
  theme_minimal()
pLI_Hubs_DF_Boxplot$values %>% hist(breaks = 100)











```



